---
title: "Homework1"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

1.  In R, use the "set.seed" function to set the random seed to "8875",
    then get 10 random numbers from an exponential distribution. What
    are the mean and median of these numbers? • Note: The values you get
    will only match the ones I'm expecting the first time you do a
    random draw after setting the seed, so if you have to redo the draw
    be sure to set the seed again, otherwise your answer will be
    incorrect.

Ans: The mean and median of the random 10 number drawn from the
exponential distribution is 0.88 and 0.85 respectively.

```{r}
set.seed(8875)
random_num<- rexp(10)
mean_exp<- mean(random_num)
median_exp<- median(random_num)
```

2.With the following matrices:

```{r}
A<- matrix(c(14,2,-6,0,-11,5), nrow = 3)
B<- matrix(c(-5,6,21,0,-3,1),nrow = 2)
AB= A%*%B
BA= B%*%A
```

3.  You are doing a linear regression to look at the effect of different
    light qualities on different varieties of clover. Growth rate is
    your dependent variable, and your two independent variables are
    clover variety (4 types) and light quality (3 types) , both of which
    are categorical variables. You have 60 observations total, so what
    will be the dimensions of your design matrix? • Note: For those with
    prior regression experience, we are not looking at any interactions
    here, just main effects.

Ans: Since we have 60 observation means the number of rows will be 60
and the two categorical variables clover variety has 4 types and light
quality has 3 types we end up having 3 and 2 factor level so at the end
the design matrix X will have the dimension of 60 * 6( intercept ,
clover variety(4-1), light quality(3-1)).

4.  You have the following data table looking at apple fruit set with
    different amounts of inputs: Solve the linear regression for this
    data using matrix algebra (like we did in class), treating Nitrogen,
    Phosphorus, and Water as quantitative variables (not factors). Show
    the following results: 
    • (X^TX)-1 \\
    • beta  \\
    • error \\
    
```{r}
#creating dataframe in R from the image 
fruit_data<- data.frame(fruitset= c(4,5.5,3,6.7,5.3,6.2),nitrogen=c(5,5,5,15,15,15), phosphorus=c(10,20,10,20,10,20),water=c(30,30,60,60,100,100))
#creating matrix for Y and X 
Y<- matrix(c(4,5.5,3,6.7,5.3,6.2),nrow = 6)
X<- matrix(c(1,1,1,1,1,1,5,5,5,15,15,15,10,20,10,20,10,20,30,30,60,60,100,100 ),nrow = 6)

#computing (X^TX)-1
Xtran<- t(X)
XtransX <-Xtran%*%X
a<- solve(XtransX) #ans for (X^TX)-1

#computing beta i.e coefficint from the equation first we will calculate X^trans * Y
XtransY<- Xtran %*% Y
beta<- a %*% XtransY

#calculating error
e<- Y- X%*%beta
sum(e^2) #0.42

```

5.	Repeat the above, but this time only include a single independent variable at a time. (That is, solve 3 models: one with Nitrogen, one with Phosphorus, and one with Water.) Which of these three models best explains the data? How do you know?
•	Note: you do not have to write down , or the like, just which is the best model and why.

Ans: Among the three model using each explanatory variable only at a time, the model using phosphorus as an explanatory variable best explain the data as it has lower sums of squared error meaning that the predicted value of the response i.e., fruit set by the model and actual fruit set recorded is close and thus the error/residual which is also known as unexplained variability is minimum when we fit the phosphorus model. 

```{r}
#model with nitrogen only
#creating matrix for Y and X 
Y<- matrix(c(4,5.5,3,6.7,5.3,6.2),nrow = 6)
X<- matrix(c(1,1,1,1,1,1,5,5,5,15,15,15 ),nrow = 6)
#computing (X^TX)-1
Xtran<- t(X)
XtransX <-Xtran%*%X
a<- solve(XtransX) #ans for (X^TX)-1

#computing beta i.e coefficint from the equation first we will calculate X^trans * Y
XtransY<- Xtran %*% Y
beta<- a %*% XtransY

#calculating error
e<- Y- X%*%beta
sum(e^2) #4.17

#using phosphorus only as an explanatory variable
#creating matrix for Y and X 
Y<- matrix(c(4,5.5,3,6.7,5.3,6.2),nrow = 6)
X<- matrix(c(1,1,1,1,1,1,10,20,10,20,10,20),nrow = 6)
#computing (X^TX)-1
Xtran<- t(X)
XtransX <-Xtran%*%X
a<- solve(XtransX) #ans for (X^TX)-1

#computing beta i.e coefficint from the equation first we will calculate X^trans * Y
XtransY<- Xtran %*% Y
beta<- a %*% XtransY

#calculating error
e<- Y- X%*%beta
sum(e^2) #3.386

#model using only water as an explanatory variable
#creating matrix for Y and X 
Y<- matrix(c(4,5.5,3,6.7,5.3,6.2),nrow = 6)
X<- matrix(c(1,1,1,1,1,1,30,30,60,60,100,100),nrow = 6)
#computing (X^TX)-1
Xtran<- t(X)
XtransX <-Xtran%*%X
a<- solve(XtransX) #ans for (X^TX)-1

#computing beta i.e coefficint from the equation first we will calculate X^trans * Y
XtransY<- Xtran %*% Y
beta<- a %*% XtransY

#calculating error
e<- Y- X%*%beta
sum(e^2) #8.5179


```



9.	You are running a GWAS for quantitative disease resistance in rice seedlings. You have a diversity panel of 580 inbred rice lines and microarray data on each of them at 120,000 sites. You filter the sites to remove any that have strong linkage between them, leaving you with 24,500 essentially independent sites. After running a GWAS, the most significant marker has a p-value of 6.82 * 10-34. What would its p-value be after performing a Bonferroni correction?<br>

Ans: The p-value after bonferroni correction will be 1.6709e-29. Basically the p-value we get here is by multiplying the p-value (6.82e-34 * no.of test(24,500)).

10.	You decide that a Bonferroni correction for the above question is too stringent and so do a false discovery rate correction instead. At FDR < 0.05, you get 124 SNPs significantly associated with your phenotypes. Assuming everything worked properly, how many of these are probably real?<br>

Ans: The real significant snps will be around 117. As the FDR correction gives us 6 false hit i.e fdr < 0.05 means 5% of 124 are our false hit.


